---
title: "Covid91 vaccine study Final2024"
author: "Luis Resendez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction
A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbiage when referencing p-values in conclusions.



$H_0$: The Covid-19 vaccine (drug) is not effective among the subgroup users.


$H_A$: The Covid-19 vaccine (drug) is effective non subgroup users.


### Males
$H_0$: The vaccine is not effective in preventing Covid-19 in males.


$H_A$: The vaccine is effective in preventing Covid-19 in males. 


Data for this study came from a FinalData file provided by Dr. James Brownlow and other sources. 

## Methods


The response method used is Cat~Cat. Data used was based on data from FinalData. We have to subset the data to include males. For graphical displays we are using: bar chart, numerical display and inferential results. 


## Results





```{r}
maleData = filter (FinalData, sex == "M")
```


## Descriptive Results
### Graphical Descriptive Results


```{r}
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData, type="percent")
```




The graph shows that among the males who got Covid-19, there was 30% who received the vaccine and 70% who got the placebo. It also shows that there was around a 10% difference in those who were OK, that received both the placebo and the vaccine.  


### Numerical Descriptive Results


```{r}
table1 <- xtabs(~infected + treatment, data=maleData)
rowPerc(table1)
colPerc(table1)
```

The table shows that the people who took the vaccine and got Covid-19 there was 2 times the difference. The males who took the placebo had a 70% chance of not being infected and 30% less likely for the people who took the vaccine. Among those who were ok and took the vaccine, there was no difference. 

The table shows that the people who took the vaccine and got Covid-19 there was 2 times the difference. There was 70% males who took the placebo that got Covid-19 and 30% of males who recievied the vaccine that got infected with Covid-19. Among those who were OK and took the vaccine or the placebo, there was no difference. 

## Inferential Results


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



The p-value is 0.00000000000000022. is so small that we failed to reject the null hypothesis. The vaccine is not effective in preventing Covid-19 among males. On the Fisher's Exact Test there is a p-value 0.00000000000000022 meaning there is a difference in getting Covid-19 and being a male.



## Females


$H_0$: The vaccine is not effective in preventing Covid-19 in females.


$H_A$: The vaccine is effective in preventing Covid-19 in females. 


The data for this study came from a FinalData file, provided by Dr. James Brownlow.


## Methods


The response method used is Cat~Cat. Data used was based on data from FinalData. We have to subset the data to include females. For graphical displays we are using: bar chart, numerical display and inferential results.


## Results



```{r}
femaleData = filter(FinalData, sex == "F")
```


## Descriptive Results


### Graphical Descriptive Results



```{r}
barchartGC(~infected+treatment,data=femaleData)
barchartGC(~infected+treatment,data=femaleData, type="percent")
```



The graph shows that there was a difference in those who got Covid-19. There was 40% females who did get the vaccine and 60% females who received the placebo. There was no significant difference in those who were OK that received either the placebo or the vaccine. 


### Numerical Descriptive Results



```{r}
table2 <- xtabs(~infected + treatment, data=femaleData)
rowPerc(table2)
colPerc(table2)
```

In this chart, it shows that there was a 29.86% difference in those who got Covid-19. There was 35.07% of people who received the vaccine and 64.93% of people who got the placebo. For the people who were OK, there was only about a 0.66% difference.

## Infrenital Results


```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```
The p-value from the Chi-squared test is  0.0001914, which is small, therefore the vaccine is not effective in preventing Covid-19. On the Fisher's Exact Test there is a p-value 0.0001246 meaning there is a difference in getting Covid-19 and being a female. We fail to reject the null hypothesis because females can still get Covid-19 with the vaccine or the placebo.



## LGBTQ



$H_0$: The vaccine is not effective in preventing Covid-19 in the LGBTQ community 


$H_A$: The vaccine is effective in preventing Covid-19 in the LGBTQ community



The data for this study came from a FinalData file, provided by Dr. James Brownlow.


## Methods


The method used is Cat~Cat. Data used was based on data from FinalData. We have to subset the data to include LGBTQ community. For graphical displays we are using: bar chart, numerical display and inferential results.


## Results


```{r}
LGBTQData = filter(FinalData, LGBTQ == "gay")
```



## Descriptive Results



### Graphical Descriptive Results



```{r}
barchartGC(~infected+treatment,data=LGBTQData)
barchartGC(~infected+treatment,data=LGBTQData, type="percent")
```


This graph shows that the people who got Covid-19, there was 80% of people who were vaccinated and 20% that received the placebo. Among those who were OK there was no significant difference. 


### Numerical Descriptive Results



```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```


In this chart, it shows that there was a 76.80% difference in those who got Covid-19. There was 20.00% of people who received the vaccine and 80.00% of people who got the placebo. For the people who were OK, there was only about a 2.90% difference.



## Infrenital Results



```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```



The Chi-squared p-value is 0.000002266, which is too small. Therefore the vaccine is not preventable against Covid-19 among the LGBTQ. In the Fisher's Exact Test the p-value is 0.0000007269 meaning there is a small difference in getting Covid-19 and being in the LGBTQ. We fail to reject the null hypothesis because people in the LGBTQ community can still get Covid-19 with the vaccine or the placebo. 



## DrugUsers


$H_0$: The vaccine is not effective in preventing Covid-19 in drug users.


$H_A$: The vaccine is effective in preventing Covid-19 in drug users.


Data for this study came from a FinalData file provided by Dr. James Brownlow.


## Methods


The response method used is Cat~Cat. Data used was based on data from FinalData. We have to subset the data to include Drugusers. For graphical displays we are using: bar chart, numerical display and inferential results. 


## Results


```{r}
DruggiesData = filter(FinalData, DrugUser == "yes")
```


## Descriptive Results


### Graphical Descriptive Results



```{r}
barchartGC(~infected+treatment,data=DruggiesData)
barchartGC(~infected+treatment,data=DruggiesData, type="percent")
```


In this graph it shows that the people who got Covid-19, there was 88% of drug users who got the vaccine and 12% of drug users who got the placebo. There was an approximate 10% difference in those who were OK, that received the placebo or the vaccine. 


### Numerical Descriptive Results



```{r}
table4 <- xtabs(~infected + treatment, data=DruggiesData)
rowPerc(table4)
colPerc(table4)
```

In this chart it shows that there was a 3.28% difference in those who got Covid-19. There was 3.61% of people who received the vaccine and 0.33% of people who got the placebo. For the people who were OK, there was only about a 2.61% difference.



## Infrenital Results



```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```



The p-value in the Chi-square Test is 0.0000000001414 which is too small, therefore the vaccine is not able to prevent Covid-19 among drug users. In the Fisher's Exact Test the p-value is 0.000000000003261 meaning there is a small difference in getting Covid-19 and being a drug user. We fail to reject the null hypothesis because drug users are able to get Covid-19 with or without the vaccine.




# Overall Results and Conclusions



In the original question we asked if the vaccine was able to prevent Covid-19 among these subgroups: males, females, LGBTQ and drug users. 

We have found that the vaccine is ineffective at preventing Covid-19 in those subgroups.

In the original question we asked if the vaccine was able to prevent Covid-19 among these subgroups: males, females, LGBTQ and drug users. We have found that the vaccine is ineffective at preventing Covid-19 in those subgroups, therefore we fail to reject the null hypothesis. 

We asked if the vaccine was effective in preventing Covid-19 

We inquired as to whether the vaccine was successful in preventing Covid-19 inside the male subgroup. Our results finally led us to conclude that the vaccine is ineffective in preventing Covid-19 among males.For this subgroup, we fail to reject the null hypothesis. 

Since the vaccine was unable to prevent Covid-19 in the subgroup of females when the identical question was posed, we fail to reject the null hypothesis. 

The LGBTQ community was also included in the question, and our data indicates that the vaccine did not successfully prevent Covid-19, thus we fail to reject the null hypothesis. 

The same questions were asked of the drug users as well, and our data indicates that the vaccination did not prevent Covid-19, therefore we fail to reject the null hypothesis.