---
title: "Covid921 Vaccine Stats Final2024 Report"
author: "Luis Resendez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction
A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbiage when referencing p-values in conclusions.




## Males
$H_0$: The Covid921 vaccine is not effective in males

$H_A$: The Covid921 vaccine is effective in males

### Methods
treatment will be Cat
infected will be Cat
helper file we are using is
CAT~CAT

### Descriptive Results

#### Graphical Results
```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```

This graph shows that the MALES who got the placebo were more infected than the MALES that were vaccinated, but about 50% that got placebos were OK, and about 50% of MALES with the vaccine were OK.


#### Numerical Results
```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```
This shows that about 74% males who took the placebo got covid19, where about 26% of males who got the vaccines got covid19

#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The statistical analysis shows a significant difference in the rate of COVID-19 infection between the placebo and vaccine groups. The chi-square test and Fisher’s Exact Test are both 0.00000000022 indicating that the vaccine significantly reduces the likelihood of contracting COVID-19. Specifically, the odds of getting COVID-19 are about 2.82 times higher for individuals receiving the placebo compared to those receiving the vaccine, with a 95% confidence interval suggesting that the vaccine is effective in preventing COVID-19 with a high degree of statistical significance.






## Females

$H_0$: The Covid921 vaccine is not effective in females

$H_A$: The Covid921 vaccine is effective in females

### Methods
Females is CAT
Treatment is CAT
Helper File we use
CAT ~ CAT

### Descriptive Results

#### Graphical Results

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```

This graph shows that the FEMALES who got the placebo were more infected than the FEMALES that were vaccinated, but about 50% that got placebos were OK, and about 51% of FEMALES with the vaccine were OK.


#### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=females)
rowPerc(table1)
colPerc(table1)
```
The numerical results shows the data indicates that the COVID-19 vaccine is effective in reducing the likelihood of contracting COVID-19 compared to the placebo. Specifically, the odds of getting COVID-19 are reduced by approximately 46.9% for those receiving the vaccine compared to those receiving the placebo. The relative risk reduction is about 45.5%, meaning that the vaccine lowers the risk of infection by roughly 45.5% compared to the placebo.



#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Chi square test being 0.00000001914 and Fisher’s Exact Test being 0.00000001246 also confirms the statistical significance of the difference in infection rates between the vaccine and placebo groups. The odds ratio of 1.8562 means that the odds of getting COVID-19 with the placebo are approximately 1.86 times higher than with the vaccine. The 95% confidence interval suggests that the true odds ratio is between 1.49 and 2.32, reinforcing the effectiveness of the vaccine.




## LGBTQ



$H_0$: The vaccine is not effective in preventing Covid-19 in the LGBTQ community 


$H_A$: The vaccine is effective in preventing Covid-19 in the LGBTQ community



## Methods

LGBTQ is CAT
Treatment is CAT
Helper File I used was CAT ~ CAT




```{r}
LGBTQData = filter(FinalData, LGBTQ == "gay")
```



## Descriptive Results
### Graphical Results



```{r}
barchartGC(~infected+treatment,data=LGBTQData)
barchartGC(~infected+treatment,data=LGBTQData, type="percent")
```


This graph shows that the LGBTQ who got the placebo were more infected than the LGBTQ that were vaccinated, but about 45% that got placebos were OK, and about 49% of LGBTQ with the vaccine were OK.



### Numerical Descriptive Results



```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```

Numerical results indicates that the COVID-19 vaccine is effective in reducing the risk of contracting the virus compared to the placebo. The odds of getting COVID-19 with the vaccine are approximately 24% of the odds with the placebo, and the risk of infection with the vaccine is about 44.8% of the risk with the placebo. The vaccine also achieves a relative risk reduction of around 55.2%, suggesting a significant protective effect.

## Infrenital Results



```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```



The Chi-squared p-value is 0.000002266, which is too small. Therefore the vaccine is not preventable against Covid-19 among the LGBTQ. In the Fisher's Exact Test the p-value is 0.0000007269The odds of getting COVID-19 with the vaccine are approximately 0.24 times the odds of getting COVID-19 with the placebo. This means that the vaccine significantly reduces the odds of contracting COVID-19 compared to the placebo. We fail to reject the null hypothesis because people in the LGBTQ community can still get Covid-19 with the vaccine or the placebo. 



## DrugUsers


$H_0$: The vaccine is not effective in preventing Covid-19 in drug users.


$H_A$: The vaccine is effective in preventing Covid-19 in drug users.


Data for this study came from a FinalData file provided by Dr. James Brownlow.


## Methods


The response method used is Cat~Cat. Data used was based on data from FinalData. We have to subset the data to include Drugusers. For graphical displays we are using: bar chart, numerical display and inferential results. 





```{r}
DruggiesData = filter(FinalData, DrugUser == "yes")
```


## Descriptive Results
### Graphical Results



```{r}
barchartGC(~infected+treatment,data=DruggiesData)
barchartGC(~infected+treatment,data=DruggiesData, type="percent")
```


In this graph it shows that the people who got Covid-19, there was 88% of drug users who got the vaccine and 12% of drug users who got the placebo. There was an approximate 10% difference in those who were OK, that received the placebo or the vaccine. 


### Numerical Descriptive Results



```{r}
table4 <- xtabs(~infected + treatment, data=DruggiesData)
rowPerc(table4)
colPerc(table4)
```

In this chart it shows that there was a 3.28% difference in those who got Covid-19. There was 3.61% of people who received the vaccine and 0.33% of people who got the placebo. For the people who were OK, there was only about a 2.61% difference.



## Infrenital Results



```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```



The p-value in the Chi-square Test is 0.0000000001414 which is too small, therefore the vaccine is not able to prevent Covid-19 among drug users. In the Fisher's Exact Test the p-value is 0.000000000003261 meaning there is a small difference in getting Covid-19 and being a drug user. We fail to reject the null hypothesis because drug users are able to get Covid-19 with or without the vaccine.




# Overall Results and Conclusions



### Males


Hypothesis Testing: We tested whether the vaccine is effective in preventing Covid-19 among males.
Results: The Chi-squared and Fisher's Exact tests showed extremely low p-values (e.g.,<0.00000000000000022
<0.00000000000000022), indicating that the vaccine did not effectively prevent Covid-19 among males. The data suggests no significant difference in Covid-19 rates between those who received the vaccine and those who received the placebo.


### Females


Hypothesis Testing: Similar to males, we tested whether the vaccine is effective in preventing Covid-19 among females.
Results: The p-values from Chi-squared and Fisher's Exact tests were 
0.0001914
0.0001914 and 
0.0001246
0.0001246 respectively, suggesting that the vaccine was not effective in preventing Covid-19 among females. There was no significant difference in Covid-19 rates between those who received the vaccine and those who received the placebo.


### LGBTQ


Hypothesis Testing:We tested whether the vaccine is effective in preventing Covid-19 among the LGBTQ community.
Results: Both Chi-squared and Fisher's Exact tests yielded very low p-values (e.g., 0.000002266
0.000002266 and 
0.0000007269
0.0000007269), indicating that the vaccine was not effective in preventing Covid-19 among the LGBTQ community. The data showed no significant difference in Covid-19 rates between vaccinated individuals and those who received the placebo.


### Drug Users


Hypothesis Testing: We tested whether the vaccine is effective in preventing Covid-19 among drug users.
Results: The Chi-squared and Fisher's Exact tests resulted in extremely low p-values (e.g., 0.0000000001414
0.0000000001414 and 
0.000000000003261
0.000000000003261), indicating that the vaccine did not effectively prevent Covid-19 among drug users. There was no significant difference in Covid-19 rates between those who received the vaccine and those who received the placebo.


### Conclusion


Based on the results across all subgroups (males, females, LGBTQ, and drug users), the data strongly suggests that the Covid-19 vaccine (Covid921) tested in these clinical trials did not effectively prevent Covid-19 compared to the placebo. Therefore, it is recommended that individuals within these subgroups consider other preventive measures or vaccines if available, given the ineffectiveness observed in this study.

This structured approach using hypothesis testing, descriptive statistics, and graphical representation provides a clear assessment of the vaccine's effectiveness across diverse population subgroups, supporting informed decision-making and further research directions.






