---
title: "Covid91 vaccine study Final2024"
author: "Luis Resendez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


## Males

$H_0$:The Covid921 Vaccine is not effective in males.


$H_A$: The Covid921 Vaccine is effective in males.




### Methods 

Treatment is cat 

Sex is cat 

helper file we are using is 

CAT~CAT 


### Descriptive Results

Based on the provided data, the count table and proportion table will provide numerical summaries of the counts and proportions of infected and non-infected individuals within each treatment group (placebo and vaccine).


#### Graphical results

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```


The first barchart shows the counts of infected and non-infected individuals for each treatment group (vaccine and placebo). This helps visualize the distribution of infection status within each treatment group.

The second barchart shows the same data but in percentage terms, showing the proportion of infected individuals within each treatment group.


#### Numerical


```{r}
# Count of infected and non-infected individuals by treatment
count_table <- table(males$treatment, males$infected)

# Proportion of infected individuals by treatment
prop_table <- prop.table(count_table, 1)

count_table
prop_table

```

The numerical results lists the counts and proportions of individuals who were infected with Covid19 "(covid19)" and those who were not infected "(OK)" within each treatment group "(placebo and vaccine)".

This table indicates that out of the individuals who received the placebo, 262 were infected with Covid19, and 9021 were not infected. Among those who received the vaccine, 93 were infected with Covid19, and 9016 were not infected.


The lower portion shows that for the placebo group, approximately 2.82% of the individuals were infected with Covid19, and 97.18% were not infected. In the vaccine group, approximately 1.02% of the individuals were infected with Covid19, and 98.98% were not infected.

These numerical results suggest that the vaccine group had a lower proportion of individuals infected with Covid19 compared to the placebo group, indicating a potential effectiveness of the vaccine in reducing the infection rate.

Summary
Placebo Group: 2.82% infection rate (262 out of 9283).
Vaccine Group: 1.02% infection rate (93 out of 9109).
The descriptive numerical results show a lower infection rate in the vaccine group compared to the placebo group, suggesting that the Covid921 vaccine might be effective in reducing the rate of Covid19 infections. Further inferential analysis will confirm the statistical significance of this observation.

#### Inferential Results

To test the hypothesis, we could use a chi-squared test to determine if there is a significant difference in infection rates between the vaccine and placebo groups:


```{r}
chisq_test <- chisq.test(count_table)
chisq_test

```
The chi-squared test results will include a p-value, which helps determine whether the observed differences are statistically significant.

#### Conclusion

Based on the descriptive and inferential results:

Since the p-value is extremely low, there is very strong statistical evidence to suggest that the Covid921 vaccine is effective in males. This supports the recommendation for males to take the Covid921 vaccine to reduce their risk of Covid19 infection.



## Females

$H_0$:The Covid921 Vaccine is not effective in females.


$H_A$: The Covid921 Vaccine is effective in females.


### Methods

### Descriptive Results

#### Graphical results

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```


#### Numerical

#### Inferential Results

#### Conclusion

Based on the statistical analysis, if the p-value < 0.05, we reject $H_0$ and conclude that the Covid921 vaccine is effective in females. Otherwise, we fail to reject $H_0$.



## LGBTQ

$H_0$:The Covid921 Vaccine is not effective in LGBTQ Community.


$H_A$: The Covid921 Vaccine is effective in LGBTQ Community.


### Methods

### Descriptive Results

#### Graphical results

#### Numerical

#### Inferential Results

#### Conclusion





## Drug Users 

$H_0$:The Covid921 Vaccine is not effective in Drug users.


$H_A$: The Covid921 Vaccine is effective in Drug users.


### Methods
The result is a categorical variable  

### Descriptive Results

#### Graphical results

```{r}
DrugUser=filter(FinalData,FinalData$DrugUser=="Yes")
barchartGC(~treatment + infected,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser,type="percent")

```


#### Numerical

#### Inferential Results

#### Conclusion



# Overall Results and Conclusions