---
title: "Covid921 Vaccine Stats Final2024 Report"
author: "Luis Resendez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction


A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbiage when referencing p-values in conclusions.




## Males


$H_0$: The Covid921 vaccine is not effective in males

$H_A$: The Covid921 vaccine is effective in males



### Methods





Treatment will be Cat
Male will be Cat
helper file we are using is
CAT~CAT

### Descriptive Results

#### Graphical Results


```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```



This graph shows that the MALES who got the placebo were more infected than the MALES that were vaccinated, but about 50% that got placebos were OK, and about 50% of MALES with the vaccine were OK.


#### Numerical Results
```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```


This shows that about 74% males who took the placebo got covid19, where about 26% of males who got the vaccines got covid19



#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The statistical analysis shows a significant difference in the rate of COVID-19 infection between the placebo and vaccine groups. The chi-square test and Fisher’s Exact Test are both 0.00000000022 indicating that the vaccine significantly reduces the likelihood of contracting COVID-19. Specifically, the odds of getting COVID-19 are about 2.82 times higher for individuals receiving the placebo compared to those receiving the vaccine, with a 95% confidence interval suggesting that the vaccine is effective in preventing COVID-19 with a high degree of statistical significance.






## Females

$H_0$: The Covid921 vaccine is not effective in females

$H_A$: The Covid921 vaccine is effective in females

### Methods


Female is CAT
Treatment is CAT
Helper File we use
CAT ~ CAT

### Descriptive Results

#### Graphical Results

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```



This graph shows that the FEMALES who got the placebo were more infected than the FEMALES that were vaccinated, but about 50% that got placebos were OK, and about 51% of FEMALES with the vaccine were OK.


#### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=females)
rowPerc(table1)
colPerc(table1)
```



The numerical results shows the data indicates that the COVID-19 vaccine is effective in reducing the likelihood of contracting COVID-19 compared to the placebo. Specifically, the odds of getting COVID-19 are reduced by approximately 46.9% for those receiving the vaccine compared to those receiving the placebo. The relative risk reduction is about 45.5%, meaning that the vaccine lowers the risk of infection by roughly 45.5% compared to the placebo.



#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Chi square test being 0.00000001914 and Fisher’s Exact Test being 0.00000001246 also confirms the statistical significance of the difference in infection rates between the vaccine and placebo groups. The odds ratio of 1.8562 means that the odds of getting COVID-19 with the placebo are approximately 1.86 times higher than with the vaccine. The 95% confidence interval suggests that the true odds ratio is between 1.49 and 2.32, reinforcing the effectiveness of the vaccine.




## LGBTQ



$H_0$: The vaccine is not effective in preventing Covid-19 in the LGBTQ community 


$H_A$: The vaccine is effective in preventing Covid-19 in the LGBTQ community



## Methods

LGBTQ is CAT
Treatment is CAT
Helper File I used was CAT ~ CAT




```{r}
LGBTQData = filter(FinalData, LGBTQ == "gay")
```



## Descriptive Results
### Graphical Results



```{r}
barchartGC(~infected+treatment,data=LGBTQData)
barchartGC(~infected+treatment,data=LGBTQData, type="percent")
```


This graph shows that the LGBTQ who got the placebo were more infected than the LGBTQ that were vaccinated, but about 45% that got placebos were OK, and about 49% of LGBTQ with the vaccine were OK.



### Numerical Descriptive Results



```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```

Numerical results indicates that the COVID-19 vaccine is effective in reducing the risk of contracting the virus compared to the placebo. The odds of getting COVID-19 with the vaccine are approximately 24% of the odds with the placebo, and the risk of infection with the vaccine is about 44.8% of the risk with the placebo. The vaccine also achieves a relative risk reduction of around 55.2%, suggesting a significant protective effect.

## Infrenital Results



```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```



The Chi-squared p-value is 0.000002266, which is too small. Therefore the vaccine is not preventable against Covid-19 among the LGBTQ. In the Fisher's Exact Test the p-value is 0.0000007269The odds of getting COVID-19 with the vaccine are approximately 0.24 times the odds of getting COVID-19 with the placebo. This means that the vaccine significantly reduces the odds of contracting COVID-19 compared to the placebo. We fail to reject the null hypothesis because people in the LGBTQ community can still get Covid-19 with the vaccine or the placebo. 



## DrugUsers


$H_0$: The vaccine is not effective in preventing Covid-19 in drug users.


$H_A$: The vaccine is effective in preventing Covid-19 in drug users.



## Methods

DrugUser is CAT
Treatment is CAT
Helper File I used was CAT ~ CAT




```{r}
DruggiesData = filter(FinalData, DrugUser == "yes")
```


## Descriptive Results
### Graphical Results



```{r}
barchartGC(~infected+treatment,data=DruggiesData)
barchartGC(~infected+treatment,data=DruggiesData, type="percent")
```


In this graph it shows that the Drug Users that got the placebo are less likely to get covid19 then those who got the vaccine. But both were about 50% ok.

### Numerical Descriptive Results



```{r}
table4 <- xtabs(~infected + treatment, data=DruggiesData)
rowPerc(table4)
colPerc(table4)
```

In this data it shows that there was a 3.28% difference in those who got Covid-19. There was 3.61% of people who received the vaccine and 0.33% of people who got the placebo. For the people who were OK, there was only about a 2.61% difference. Although, This analysis shows an unusual result where the vaccine appears to increase the risk of infection compared to the placebo, which is contrary to expectations. This could suggest either an anomaly in the data or a specific context of the study that should be reviewed. In a typical scenario, a vaccine should reduce the risk of COVID-19, not increase it.



## Infrenital Results



```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```



The p-value in the Chi-square Test is 0.0000000001414 which is too small, therefore the vaccine is not able to prevent Covid-19 among drug users. In the Fisher's Exact Test the p-value is 0.000000000003261 meaning there is a small difference in getting Covid-19 and being a drug user. We fail to reject the null hypothesis because drug users are able to get Covid-19 with or without the vaccine.




# Overall Results and Conclusions



## Males


Descriptive Results and Graphical Results graphs shows that males who received the placebo had a higher rate of COVID-19 infection compared to those who received the vaccine. About 50% of males in both groups were "OK," indicating that the placebo group had a higher infection rate.

Numerical Results show approximately 74% of males who took the placebo got COVID-19, while about 26% of males who received the vaccine got COVID-19.

Inferential Results show that statistical tests revealed a significant difference in infection rates between the placebo and vaccine groups:

Chi-Square Test p-value: 2.2

Fisher’s Exact Test p-value: 2.2

Odds Ratio: 2.82 (95% CI: [2.21, 3.61])

So the vaccine is highly effective for males. The analysis indicates that males who received the vaccine had significantly lower odds of contracting COVID-19 compared to those who received the placebo. The odds of getting COVID-19 were about 2.82 times higher for those receiving the placebo compared to those receiving the vaccine. This suggests that the vaccine offers strong protection against COVID-19 for males. So we reject the null hypothesis. The Covid921 vaccine is effective in preventing COVID-19 in males.

## Females


Descriptive Results and Graphical Results graphs shows that females who received the placebo had a higher rate of COVID-19 infection compared to those who received the vaccine. Both groups had about a 50% "OK" rate, with a slight increase in the "OK" percentage for the vaccine group.

Numerical Results show the odds of getting COVID-19 were reduced by approximately 46.9% for those receiving the vaccine compared to those receiving the placebo. The relative risk reduction was about 45.5%.

Inferential Results
The statistical tests confirmed a significant difference in infection rates:

Chi-Square Test p-value: 1.914

 
Fisher’s Exact Test p-value: 1.246

 
Odds Ratio: 1.86 (95% CI: [1.49, 2.32])

So the vaccine is effective for females. The analysis demonstrates that the COVID-19 vaccine significantly reduces the risk of infection in females. Specifically, the odds of getting COVID-19 were about 1.86 times higher for those receiving the placebo compared to those receiving the vaccine, indicating that the vaccine is effective in preventing COVID-19 for females as well. So we reject the null hypothesis. The Covid921 vaccine is effective in preventing COVID-19 in females.



## LGBTQ Community

Descriptive Results and Graphical Results show that LGBTQ individuals who received the placebo had a higher rate of COVID-19 infection compared to those who received the vaccine. Both groups had a somewhat lower "OK" rate compared to the male and female groups.

Numerical Results Show the odds of getting COVID-19 with the vaccine were approximately 24% of the odds with the placebo. The risk of infection with the vaccine was about 44.8% of the risk with the placebo, indicating a significant protective effect.

Inferential Results
The statistical tests showed a significant difference in infection rates:

Chi-Square Test p-value: 2.266

 
Fisher’s Exact Test p-value: 7.269

Odds Ratio: 0.24 (95% CI: [0.17, 0.33])

So the vaccine is effective for the LGBTQ community. The analysis indicates that the vaccine significantly reduces the risk of COVID-19 infection among LGBTQ individuals. The odds of contracting COVID-19 were about 0.24 times the odds for those receiving the placebo, demonstrating that the vaccine provides significant protection. So we reject the null hypothesis. The Covid921 vaccine is effective in preventing COVID-19 in the LGBTQ community.

## Drug Users

Descriptive Results and Graphical Results show that drug users who received the placebo had a lower rate of COVID-19 infection compared to those who received the vaccine. Both groups had a similar "OK" rate, with the vaccine group showing a slightly higher infection rate.

Numerical Results analysis shows a 3.28% difference in the rate of COVID-19 between the vaccine and placebo groups, with the vaccine appearing to increase the risk of infection, contrary to expectations.

Inferential Results
The statistical tests indicated significant results, but with an unusual outcome:

Chi-Square Test p-value: 1.414

 
Fisher’s Exact Test p-value: 3.261

Odds Ratio: 4.27 (95% CI: [2.22, 8.88])

So the vaccine does not appear to be effective for drug users. The data reveals an unexpected result where the vaccine group had a higher infection rate compared to the placebo group. This suggests there may be an anomaly in the data or a unique context affecting drug users, which warrants further investigation. In this case, the vaccine did not provide the expected protective effect and might have increased the risk of COVID-19. So we fail to reject the null hypothesis.

## Overall Conclusion

In summary, the COVID-19 vaccine demonstrated effectiveness across most groups analyzed, showing significant reductions in COVID-19 infection rates among males, females, and the LGBTQ community. For males and females, the vaccine significantly reduced the odds of contracting COVID-19 compared to the placebo. However, the data for the LGBTQ community also confirmed the vaccine's effectiveness. In contrast, the analysis for drug users revealed an unusual outcome where the vaccine appeared to increase the risk of COVID-19 infection, which suggests the need for further review of this specific group.


