---
title: "Covid921 Vaccine Stats Final2024 Report"
author: "Luis Resendez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction
A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbiage when referencing p-values in conclusions.




## Males
$H_0$: The Covid921 vaccine is not effective in males

$H_A$: The Covid921 vaccine is effective in males

### Methods
treatment will be Cat
infected will be Cat
helper file we are using is
CAT~CAT

### Descriptive Results

#### Graphical Results
```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```

This chart shows that males who got the actual vaccines are less likely to get Covid921,
than those who got the placebo vaccine.


#### Numerical Results
```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```
This shows that about 74% males who took the placebo got covid19, where about 26% of males who got the vaccines got covid19

#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
When running the Chi-squared we get a pvalue of .00000000000000022 which is less than 0.05. This means that we reject the null hypothesis that Covid921 vaccine is not effective in males

When running the Fisher Exact test we have .00000000000000022 which is also less than 0.05.

The odds ratio would be 2.81 : 1, which means that their is a 2.81 is more likely to get when taking placebo, than those who are 1 times more likely to get covid when taking the vaccines.

#### Conclusion

Our question was to find if the Covid19 is effected in males or not effective in males. When doing our graphical result it shows that males who got the covid vaccines are less likely to get covid then those who got Placebo. When indicating the numerical results, it shows that about 74% males who took the placebo got covid19, where about 26% of males who got the vaccines got covid19. We ran the Chi-square and got the Pvalue of .00000000000000022 which is less than .05. This indicates that we reject the null hypothesis. When doing the fisher test we came out with the same hypothesis as the chi-sqr. There for we reject the null hypothesis again. We conclude that taking the covid19 vaccine is effective in males.




## Females

$H_0$: The Covid921 vaccine is not effective in females

$H_A$: The Covid921 vaccine is effective in females

### Methods
Females is CAT
Treatment is CAT
Helper File we use
CAT ~ CAT

### Descriptive Results

#### Graphical Results

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```

This chart shows that females who got the actual vaccines are less likely to get Covid921,
than those who got the placebo vaccine.

#### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=females)
rowPerc(table1)
colPerc(table1)
```
When doing the numerical results it shows that about 65% of women who got placebos are more likely to get Covid19 than those who actually got the vaccines which was about 35% of women.

#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


## LGBTQ



$H_0$: The vaccine is not effective in preventing Covid-19 in the LGBTQ community 


$H_A$: The vaccine is effective in preventing Covid-19 in the LGBTQ community



The data for this study came from a FinalData file, provided by Dr. James Brownlow.


## Methods


The method used is Cat~Cat. Data used was based on data from FinalData. We have to subset the data to include LGBTQ community. For graphical displays we are using: bar chart, numerical display and inferential results.





```{r}
LGBTQData = filter(FinalData, LGBTQ == "gay")
```



## Descriptive Results
### Graphical Results



```{r}
barchartGC(~infected+treatment,data=LGBTQData)
barchartGC(~infected+treatment,data=LGBTQData, type="percent")
```


This graph shows that the people who got Covid-19, there was 80% of people who were vaccinated and 20% that received the placebo. Among those who were OK there was no significant difference. 


### Numerical Descriptive Results



```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```


In this chart, it shows that there was a 76.80% difference in those who got Covid-19. There was 20.00% of people who received the vaccine and 80.00% of people who got the placebo. For the people who were OK, there was only about a 2.90% difference.



## Infrenital Results



```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```



The Chi-squared p-value is 0.000002266, which is too small. Therefore the vaccine is not preventable against Covid-19 among the LGBTQ. In the Fisher's Exact Test the p-value is 0.0000007269 meaning there is a small difference in getting Covid-19 and being in the LGBTQ. We fail to reject the null hypothesis because people in the LGBTQ community can still get Covid-19 with the vaccine or the placebo. 



## DrugUsers


$H_0$: The vaccine is not effective in preventing Covid-19 in drug users.


$H_A$: The vaccine is effective in preventing Covid-19 in drug users.


Data for this study came from a FinalData file provided by Dr. James Brownlow.


## Methods


The response method used is Cat~Cat. Data used was based on data from FinalData. We have to subset the data to include Drugusers. For graphical displays we are using: bar chart, numerical display and inferential results. 





```{r}
DruggiesData = filter(FinalData, DrugUser == "yes")
```


## Descriptive Results
### Graphical Results



```{r}
barchartGC(~infected+treatment,data=DruggiesData)
barchartGC(~infected+treatment,data=DruggiesData, type="percent")
```


In this graph it shows that the people who got Covid-19, there was 88% of drug users who got the vaccine and 12% of drug users who got the placebo. There was an approximate 10% difference in those who were OK, that received the placebo or the vaccine. 


### Numerical Descriptive Results



```{r}
table4 <- xtabs(~infected + treatment, data=DruggiesData)
rowPerc(table4)
colPerc(table4)
```

In this chart it shows that there was a 3.28% difference in those who got Covid-19. There was 3.61% of people who received the vaccine and 0.33% of people who got the placebo. For the people who were OK, there was only about a 2.61% difference.



## Infrenital Results



```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```



The p-value in the Chi-square Test is 0.0000000001414 which is too small, therefore the vaccine is not able to prevent Covid-19 among drug users. In the Fisher's Exact Test the p-value is 0.000000000003261 meaning there is a small difference in getting Covid-19 and being a drug user. We fail to reject the null hypothesis because drug users are able to get Covid-19 with or without the vaccine.




# Overall Results and Conclusions



### Males


Hypothesis Testing: We tested whether the vaccine is effective in preventing Covid-19 among males.
Results: The Chi-squared and Fisher's Exact tests showed extremely low p-values (e.g.,<0.00000000000000022
<0.00000000000000022), indicating that the vaccine did not effectively prevent Covid-19 among males. The data suggests no significant difference in Covid-19 rates between those who received the vaccine and those who received the placebo.


### Females


Hypothesis Testing: Similar to males, we tested whether the vaccine is effective in preventing Covid-19 among females.
Results: The p-values from Chi-squared and Fisher's Exact tests were 
0.0001914
0.0001914 and 
0.0001246
0.0001246 respectively, suggesting that the vaccine was not effective in preventing Covid-19 among females. There was no significant difference in Covid-19 rates between those who received the vaccine and those who received the placebo.


### LGBTQ


Hypothesis Testing:We tested whether the vaccine is effective in preventing Covid-19 among the LGBTQ community.
Results: Both Chi-squared and Fisher's Exact tests yielded very low p-values (e.g., 0.000002266
0.000002266 and 
0.0000007269
0.0000007269), indicating that the vaccine was not effective in preventing Covid-19 among the LGBTQ community. The data showed no significant difference in Covid-19 rates between vaccinated individuals and those who received the placebo.


### Drug Users


Hypothesis Testing: We tested whether the vaccine is effective in preventing Covid-19 among drug users.
Results: The Chi-squared and Fisher's Exact tests resulted in extremely low p-values (e.g., 0.0000000001414
0.0000000001414 and 
0.000000000003261
0.000000000003261), indicating that the vaccine did not effectively prevent Covid-19 among drug users. There was no significant difference in Covid-19 rates between those who received the vaccine and those who received the placebo.


### Conclusion


Based on the results across all subgroups (males, females, LGBTQ, and drug users), the data strongly suggests that the Covid-19 vaccine (Covid921) tested in these clinical trials did not effectively prevent Covid-19 compared to the placebo. Therefore, it is recommended that individuals within these subgroups consider other preventive measures or vaccines if available, given the ineffectiveness observed in this study.

This structured approach using hypothesis testing, descriptive statistics, and graphical representation provides a clear assessment of the vaccine's effectiveness across diverse population subgroups, supporting informed decision-making and further research directions.






